We'll start by loading the necessary packages and data sets
```{r, message=FALSE, echo = TRUE}
library("dplyr")
library("ggplot2")
data(okcupid, package = "jrTidyverse")
```

## Summarising the data

In this section, we will gradually chain the commands together. We'll 
start things off, by calculating the average income

```{r, echo = TRUE}
okcupid %>% 
  summarise(ave_income = mean(income))
```

1. Alter the above command to calculate the median income (as well as the mean).

```{r}
okcupid %>% 
  summarise(ave_income = mean(income),
            med_income = median(income))
```


2. Using `filter()` and `is.na()` remove people who skipped the `pet` question before calculating the average income.

```{r}
okcupid %>%
  filter(!is.na(pets)) %>%
  summarise(ave_income = mean(income))
```


3. Use the `group_by()` to calculate the average incomes conditional on the answer
to the `pets` question.

```{r}
okcupid %>%
  filter(!is.na(pets)) %>%
  group_by(pets) %>%
  summarise(ave_income = mean(income))
```


4. The `arrange()` function is used to sort a tibble, .e.g

    ```{r, eval = FALSE, echo = TRUE}
    ... %>%
      arrange(ave_income)
    ```
  will arrange the tibble from smallest to largest. Arrange the tibble from __largest__
to smallest in terms of average income.

```{r}
(df = okcupid %>%
  filter(!is.na(pets)) %>%
  group_by(pets) %>%
  summarise(ave_income = mean(income)) %>%
  arrange(desc(ave_income))
)
```

5. Using __ggplot2__ and `geom_col()` plot your results. Hint use `+ coord_flip()` to rotate
your plot.

```{r}
ggplot(df) + 
  geom_col(aes(x = pets, y = ave_income)) + 
  coord_flip()
```


<!-- XXX not sure if you intended to work through the above together or not but seemed weird having this by itself -->
<!-- ## All by yourself -->

<!-- 6. Calculate the average age in the OK Cupid data set, conditional on gender. This should take -->
<!-- three lines of R code. -->

<!-- ```{r} -->
<!-- okcupid %>%  -->
<!--   group_by(sex) %>% -->
<!--   summarise(ave_age = mean(age)) -->
<!-- ``` -->


## Creating columns with `mutate()`

1. The `floor()` function rounds down to the nearest integer. To round to the nearest $10$, we use the trick
    ```{r, echo=TRUE}
    floor(61/10)*10
    floor(119/10)*10
    ```
  We can use the `mutate()` function to create a new column that contains the persons
  age (to the decade), i.e. 50, 60, 70, etc.  The `mutate()` function isn't directly in the notes, but it is relatively easy to understand. It creates a new column with the given name, based on manipulation of existing columns. So we could create this new column `decade`.

    ```{r, echo = TRUE}
    okcupid %>%
      mutate(decade = floor(age/10) * 10)
    ```

2. Since this data set has high earners, use `filter()` to remove the 
top 5\% of earners. Hint: `quantile(income, probs = 0.95)` will give you the 
95\%-tile  of income

```{r}
okcupid %>%
  mutate(decade = floor(age/10)*10) %>%
  filter(income < quantile(income, probs = 0.95))
```

3. To help with plotting, convert the `decade` column into a character (hint use `mutate()`)

```{r}
(df = okcupid %>%
  mutate(decade = floor(age/10)*10) %>%
  filter(income < quantile(income, probs = 0.95)) %>%
  mutate(decade = as.character(decade))
)
```


4. Use __ggplot2__ to create boxplots of `x = decade` and `y = income`.

```{r}
ggplot(df) + 
  geom_boxplot(aes(x = decade, y = income))
```

5. Create facets by using `+ facet_wrap(~ drugs)`

```{r}
ggplot(df) + 
  geom_boxplot(aes(x = decade, y = income)) + 
  facet_wrap(~drugs)
```

