---
title: "Chapter 3 Practical"
author: "Jumping Rivers"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache= TRUE)
```

## Getting Started

First load the ggplot2 package
```{r, message = FALSE}
library("ggplot2")
```
and the OK Cupid data set
```{r}
data(okcupid, package = "jrTidyverse")
```

## Scatter plots

Let's start by creating a basic scatter plot of the heights and ages. We do this using the `geom_point()` command

```{r, fig.keep='none'}
# alpha makes the points transparent
ggplot(data = okcupid) + 
  geom_point(aes(x = age, y = height), alpha = 0.2)
```

To save typing, we're going to store the original ggplot object as a variable

```{r, fig.keep='none'}
g = ggplot(data = okcupid)
g1 = g + geom_point(aes(x = age, y = height), alpha = 0.2)
```

So now running g1 will produce the graph
```{r, fig.keep='none'}
g1
```

The arguments `x` and `y` here are called aesthetics. What do you think happens if you omit the `y` aesthetic?

For `geom_point()`, these aesthetics are required. This particular geom has other aesthetics: shape, colour, size and alpha.^[These are available for most geoms]

For instance

```{r, fig.keep='none'}
g + geom_point(aes(x = age, y = height, colour = sex))
```

Try changing using other variables (from the data set) and other aesthetics.

## Bar plots

Try `geom_bar()` with `x` and `y` variables? What happens? 

`geom_bar()` requires only one aesthetic and that is `x`. 

```{r, fig.keep='none'}
g + geom_bar(aes(x = body_type))
```

We can change the axis labels by adding two more layers using `xlab()` and `ylab()`. 

```{r, fig.keep='none'}
g + geom_bar(aes(x = body_type)) + 
  xlab("Body Type") + 
  ylab("Total")
```

Split the graph up into the two genders (hint: use `colour` and `fill`). 

```{r, fig.keep='none'}
# What happens if you only have colour or only fill?
g + geom_bar(aes(x = body_type, colour = sex, fill = sex))
```

Try a coordinate flip using `coord_flip()`

```{r, fig.keep='none'}
g + 
  geom_bar(aes(x = body_type, colour = sex, fill = sex)) + 
  coord_flip()
```

I am not too keen on how the female and male bars are displayed on top of each other. The argument to change these is `position`. The different options are listed in table \@ref(T1). The default is `stack`, e.g.  

```{r, fig.keep='none'}
g + geom_bar(aes(x = body_type, colour = sex, fill = sex), position = "stack") + 
  coord_flip()
```

Try changing the position argument - try position equal to `dodge`, `fill`, `jitter`, or `identity`. 

