We've already covered a fair bit of tidyverse stuff in the Intro to R course. Namely tibbles, **dplyr** and **ggplot2**. This chapter just serves as a general recap into what we touched upon. 

## tibbles

A tibble (or data frame) is how we store a sheet of data. A standard tibble looks like this 
```{r, echo = TRUE, results="show"}
data(example, package = "jrTidyverse")
example
```

## dplyr

**dplyr** is a package for manipulating tibbles. We covered several functions, such as `filter()` and `summarise()`
```{r, message = FALSE, warning = FALSE, echo = TRUE, results='show'}
library("dplyr")
## Give me all the rows where gender is "Male"
filter(example, gender == "Male")
## What is the average of the age variable
summarise(example, av_age = mean(age))
```

1. Filter the example data set so that the rows left have `age` > `24`. 
```{r}
filter(example, age > 24)
```

2. Filter the example data set so that the rows left have `respond` = `TRUE`. 
```{r}
filter(example, respond)
```

We can pass outputs to the first argument of the next function using the piping operator, `%>%`
```{r, echo = TRUE, results='show'}
# Give me the average age of males
example %>%
  filter(gender == "Male") %>%
  summarise(av_age = mean(age))
```

3. Grab the average age of the people who didn't respond
```{r}
example %>%
  filter(!respond) %>%
  summarise(av_age = mean(age))
```


The piping operator can be used in any functions, not just **dplyr**
```{r, echo = TRUE, results='show'}
# Pass 1:5 on the left as the first argument to mean
1:5 %>%
  mean(na.rm = TRUE)
# Explicitly pass 1:5 into the function
mean(1:5, na.rm = TRUE)
```

We can apply functions to groups within variables using `group_by()`
```{r, echo = TRUE, results='show'}
# Give me the average age of each group within gender
example %>%
  group_by(gender) %>%
  summarise(av_age = mean(age))
```

5. Grab the average age of each grouping within respond
```{r}
example %>%
  group_by(respond) %>%
  summarise(av_age = mean(age))

```

## ggplot2

**ggplot2** is a fantastic package for graphics. The `ggplot()` function creates a **ggplot2** object.
```{r,results='hide', fig.keep='none', message = FALSE, warning = FALSE, echo = TRUE}
library("ggplot2")
data(movies, package = "jrTidyverse")
ggplot(movies)
```

To add axes to this we add aesthetics 
```{r, label = "2-1", fig.margin  = TRUE, fig.cap = "Initial ggplot object", echo = TRUE, fig.keep='all', fig.height=3, fig.width=3, fig.pos="h"}
# visible in figure 2.1
g = ggplot(movies, aes(x = duration, y = rating))
g
```
Notice we can save plots as variables.

Then to add information onto the graph we use geoms
```{r, eval = FALSE, message = FALSE, warning = FALSE, echo = TRUE, results='show', fig.keep='all'}
# figure 2.2
g +
  geom_point()
ggplot(movies, aes(x = rating)) +
  geom_histogram()
ggplot(movies, aes(x = classification)) +
  geom_bar()
ggplot(movies, aes(x = classification, y = rating)) +
  geom_boxplot()
```
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height=6, fig.width=6, fig.cap = "Examples of different geoms", fig.keep='all', eval = TRUE, results="show"}
a = g +
  geom_point()
b = ggplot(movies, aes(x = rating)) +
  geom_histogram()
c = ggplot(movies, aes(x = classification)) +
  geom_bar()
d = ggplot(movies, aes(x = classification, y = rating)) +
  geom_boxplot()
gridExtra::grid.arrange(a, b, c, d, ncol = 2, nrow = 2)
```

1. Using **dplyr**, work out the average rating for each year of movies in the data set. 
```{r}
movies %>%
  group_by(year) %>%
  summarise(av_rat = mean(rating))
```

2. Use **ggplot2** to plot a line graph of the year against average rating. 
**Hint:** Use `geom_line()`
```{r}
movies %>%
  group_by(year) %>%
  summarise(av_rat = mean(rating)) %>%
  ggplot(aes(x = year, y = av_rat)) +
  geom_line()
```

3. Change the axis labels and titles to something more sensible. Remember, to change labels and titles, we use `labs()`
```{r}
movies %>%
  group_by(year) %>%
  summarise(av_rat = mean(rating)) %>%
  ggplot(aes(x = year, y = av_rat)) +
  geom_line() +
  labs(x = "Year",
       y = "Average Rating",
       title = "Average rating per year of movies")
```
